\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-02-07 paper/main.tex
Results point 1

Within DTI prediction, one can observe basic biases resulting from the
underlying datasets \citep{Pahikkala2014}. First, novel drugs are
often designed by altering non-functional components of a drug,
leading to two and more very similar drugs designed to target the same
proteins \cite{Overington2006}. This can lead to a bias when it leads
to \glqq hidden duplicates\grqq{} that can distribute among the
train--test split, resulting in a better (measured) predictive
performance than would be expected when the model is applied to
genuinely unknown interactions. Second, some proteins (which we will
denote as \textit{hub proteins}) have significantly more known
interactions with drugs than others. In the STITCH database, $5\%$
of the proteins have $40\%$ of the interactions, and similar
distributions are present in the Yamanishi and Drugbank datasets
\cite{Drugbank2007, Drugbank2017} datasets; preferentially predicting
these proteins can often increase predictive performance while again
not reflecting the actual performance when applied to a new protein
(e.g., a protein for which no interactions are known). This bias in
the number of drugs targeting certain proteins may be the result of
study bias where more \glqq valuable\grqq{} proteins have more drugs
designed to target them due to their involvement in more common
diseases (or diseases for which drugs can be more profitably
marketed).

We developed \name as a computational model to predict drug--target
interactions. Specifically, given a protein, \name will identify and
rank known drugs that likely target this protein. \name combines two 
types of features: structural information for drugs and proteins that
can be used to determine if they may physically interact, and
information about drug effects that may localize on an interaction
network.
As structural features, \name uses structural representations of drugs
from the SMILES transformer \cite{SmilesTransformer} and representations of protein
amino acid sequences from DeepGOPlus \cite{DeepGoPlus}.
\name learns representations of drug effects and protein functions
using the ontology-based machine learning method DL2Vec \cite{DL2vec2020}.
When combining ontology and molecular features with or without the graph model, we concatenate both protein features and both drugs features, before plugging them into the graph or the similarity computation.

>>>>

Within DTI prediction, there are potential biases resulting from the
underlying datasets \citep{Pahikkala2014}. First, novel drugs are
often designed by altering non-functional components of a drug,
leading to two and more very similar drugs designed to target the same
proteins \cite{Overington2006}. This can result in a bias when it
leads to hidden duplicates that can distribute among the train/test
split, resulting in a better (measured) predictive performance than
would be expected when the model is applied to identify drugs that
target a protein for which no drugs yet exist. Second, some proteins
(which we will denote as \textit{hub proteins}) have significantly
more known interactions with drugs than others. In the STITCH
database, $5\%$ of the proteins have $40\%$ of the interactions, and
similar distributions are present in the Yamanishi and Drugbank
datasets \cite{Drugbank2007, Drugbank2017} datasets; preferentially
predicting these proteins may increase predictive performance while
again not reflecting the actual performance when applied to a new
protein (e.g., a protein for which no interactions are known). These
differences in the number of drugs targeting certain proteins may be
the result of study bias where more ``valuable'' proteins have more
drugs designed to target them due to their involvement in more common
diseases (or diseases for which drugs can be more profitably
marketed).

We developed \name as a computational model to predict drug--target
interactions. Specifically, given a protein, \name will identify and
rank known drugs that likely target this protein. \name combines two
types of features: structural information for drugs and proteins that
can be used to determine if they may physically interact, and
information about drug effects that may localize on an interaction
network.  As structural features, \name uses structural
representations of drugs from the SMILES transformer
\cite{SmilesTransformer} and representations of protein amino acid
sequences from DeepGOPlus \cite{DeepGoPlus}.  \name learns
representations of drug effects and protein functions using the
ontology-based machine learning method DL2Vec \cite{DL2vec2020}.

When combining ontology and molecular features with or without the
graph model, we concatenate both protein features and both drugs
features, before plugging them into the graph or the similarity
computation.
